Robust Spectroscopic Inference with Imperfect Models

### 2014-09-19 tex/paper.tex
last call edits on full draft; ready to circulate after your changes (if any)

%\documentclass[preprint]{aastex} 
\documentclass[iop,floatfix]{emulateapj}

@@

\newcommand{\vM}{\mathsf{M}}
\newcommand{\vD}{\mathsf{D}}
\newcommand{\vR}{\mathsf{R}}
\newcommand{\vC}{\mathsf{C}}
\newcommand{\fM}{ \vec{{\bm M}}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ \vec{{\bm D}}}
\newcommand{\fDi}{D_i}
\newcommand{\fR}{ {\bm R}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}
\newcommand{\matern}{Mat\'{e}rn}
\newcommand{\HK}{$\textrm{H}_2$O-K2}

%Chebyshev coefficients RULE!
\newcommand{\cc}[2]{c_{#2}^{(#1)}} %echelle order first, then degree coefficient

%flux
\newcommand{\flam}{f_\lambda}

%\vt stands for ``vector theta''
\newcommand{\vt}{ {\bm \theta}}
\newcommand{\vT}{ {\bm \Theta}}

%\vp stands for ``vector phi''
\newcommand{\vp}{ {\bm \phi}}
\newcommand{\vP}{ {\bm \Phi}}

%\phi and \Phi are used for nuisance parameters
\newcommand{\cheb}{ \vp_{\mathsf{P}}}
\newcommand{\chebi}[1]{ \vp_{\textrm{Cheb}_{#1}}} %for signifying a specific order
\newcommand{\Cheb}{ \vP_{\textrm{Cheb}}}
\newcommand{\Chebi}[1]{ \vP_{\textrm{Cheb}_{\ne #1}}} %for specifying everything but a specific order
\newcommand{\cov}{ \vp_{\mathsf{C}}}
\newcommand{\covi}[1]{ \vp_{\textrm{cov}_{#1}}} %for signifying a specific order
\newcommand{\Cov}{ \vP_{\textrm{cov}}}
\newcommand{\Covi}[1]{ \vP_{\textrm{cov}_{\ne #1}}} %for specifying everything but a specific order

\newcommand{\allParameters}{\vT} %all parameters
\newcommand{\nuisanceParameters}{\vP} %all nuisance parameters

%kernel notation
\newcommand{\KK}{\mathcal{K}}
\newcommand{\Kglobal}{\KK^{\textrm{g}}}
\newcommand{\Klocal}{\KK^l}

@@

\shorttitle{Spectroscopic inference}
\shortauthors{Czekala et al.}

@@

%\title{A Method for the Spectroscopic Inference of Stellar Parameters}
\title{Robust Spectroscopic Inference with Imperfect Models}
\author{Ian Czekala, Sean M.~Andrews, Guillermo Torres, \& David W.~Latham}
\affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\email{iczekala@cfa.harvard.edu}

\begin{abstract}
We present a modular and extensible framework for forward-modelling spectra with synthetic models. Due to the nature of spectra, that adjacent pixels are not independent samples of the intrinsic flux distribution, it is common to result in residual spectra that exhibit some degree of correlation. Neglecting this covariance in the residual spectrum can result in a biased estimates of the parameters which describe the model spectrum. In the high signal-to noise, large spectral range limit common to stellar parameter estimation, the covariant structure of the residuals can bias the resulting parameter determinations. We present a likelihood function designed to account for this covariant structure, utilizing the machinery of Gaussian process covariance kernels to self-consistently identify and model the covariance. Specific to stellar parameter estimation, a common problem is the presence of spectral lines with an incorrect strength, often due to incorrect atomic constants or radiative treatment. We also present the development of a local covariance kernel that can track and self-consistently down-weight pathologically wrong spectra lines. Local kernels that track line weights provide a mechanism to learn corrections to spectral models, by fitting in a hierarchical manner. We demonstrate the application of this method, implemented as an open source code, to fit an echelle optical spectrum WASP-14, an F star, and a mid-infra spectrum of Gl51, an M5 star. 
\end{abstract}
\keywords{stars: fundamental parameters --- methods: data analysis --- methods: statistical}

@@

All astronomers recognize that spectroscopy offers a wealth of information that can help
characterize the properties of the observing target.  In the context of stellar astrophysics, 
spectroscopy plays many fundamental roles.  The relative strengths and widths of stellar absorption 
lines provide access to key parameters like effective temperature ($T_{\rm eff}$) and surface 
gravity ($\log g$), enabling model comparisons in the Hertzsprung-Russell diagram to estimate the 
masses and ages so crucial to understanding stellar evolution, as well as individual elemental (and 
molecular) abundances or the collective ``metallicity" (typically parameterized as [Fe/H]), 
facilitating study of the chemical hallmarks of different stellar populations.  With sufficient 
spectral resolution, the velocity content of a spectrum can convey crucial information about 
stellar rotation ($v \sin i$) and kinematics (e.g., association with a cluster or companion through 
the radial velocity, $v_z$).  While many fields benefit from the spectroscopic measurements of 
these stellar properties, there is acute interest in the exoplanet community.  There, all estimates 
of the planet properties are made {\it relative} to the host properties (e.g., the planet-to-host 
mass or radius {\it ratio} is constrained with the radial velocity or transit techniques, 
respectively).  Moreover, essential clues to the planet formation process are encapsulated in the 
dependences of planet frequency on host mass \citep[e.g.,][]{johnson07,howard10} and metallicity 
\citep[e.g.,][]{fischer05,buchhave14}.   
% SA: we ought to check with exoplanet people on these references. Can do later...

@@

There are various well-tested techniques being used to compare these models with observed spectra 
and thereby infer basic stellar parameters; here we highlight three examples to illustrate some key 
ideas.  First is a straightforward empirical approach that relies on distilling an information-rich 
subset of the data, usually in the form of spectral line equivalent widths and/or local continuum 
shapes.  A combined sequence of the ratios of these quantities are found to be especially sensitive 
to one key parameter in the models, and therefore can be used to quantify it 
\citep[e.g.,][]{gray94,reid95,rojas-ayala10,rojas-ayala12}.  This index comparison approach has 
the advantage of being trivially fast, but each index relationship is only informative over a 
limited region of parameter-space.  A second technique exploits the cross-correlation of an 
observed spectrum with a suite of model templates to identify an optimized set of parameters, 
usually with some weighting applied to specific spectral regions \citep[e.g., {\tt 
SPC};][]{buchhave12}.  In this case, the speed advantage is maintained and more data content is 
used (particularly in the spectral dimension), thereby achieving higher precision even for data 
with comparatively low sensitivity; the disadvantage is that the model quality and parameter 
inferences are assessed in an empirical, rather than probabilistic, framework.  A third approach 
employs a direct, pixel-by-pixel comparison between model and data, and is usually associated with 
a spectral synthesis back-end (e.g., {\tt MOOG}; \citealt{sneden73}; {\tt SME}; 
\citealt{valenti96}).  This technique has the benefits of parametric flexibility (e.g., one can fit 
for arbitrary abundances or structures) and a proper inference framework \citep[usually a 
least-squares approach, although increasingly in a Bayesian format;][]{shkedy07,schoenrich13}, but 
it is computationally expensive and can be biased by systematics \citep[e.g.,][]{mann13}; in 
practice, this often requires an effort focused on a pre-selected subsample of the data.

In this article, we design a flexible forward-modeling approach to the general spectroscopic 
inference problem in a Bayesian framework, building on the best aspects of the latter two methods 
highlighted above.  Employing a non-trivial covariance matrix parameterized by both global 
(stationary) and local (non-stationary) Gaussian process kernels, we are able to account for 
residual pixel-to-pixel correlations that arise when comparing observed spectra to intrinsically 
imperfect models.  This approach minimizes the potential for bias, efficiently propagates 
systematic uncertainties into the parameter inferences, and ultimately can serve as tool for using
observations to guide substantially improvements in the models themselves.  An overview of the 
methodology of this approach is provided in Section \ref{sec:method}, including the covariance 
matrix formalism.  Some tests and example applications (for a high resolution optical spectrum of 
an F star, and a medium-resolution near-infrared spectrum of a mid-M star) of the method are 
described in Section \ref{sec:examples}.  Finally, a discussion of the potential utility of the 
technique, and the possibility of extending it to develop data-driven spectral models, is provided 
in Section \ref{sec:discussion}.  \\

@@

We opt for the computationally expedient approach that employs a library of
model spectra, $\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm
eff}, \,\, \log{g}, \,\, [{\rm Fe/H}]]$.  However, it is worth noting that the
techniques developed here are applicable to {\it any} ``back-end" that
generates a model spectrum.  In our adopted approach, the model spectrum for an
arbitrary $\vt_{\ast}$ is interpolated from a spectral library,
\begin{equation} \label{eqn:interp} \flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto
\flam(\vt_{\ast}), \end{equation} where we assign the symbol $\leadsto$ as an
interpolation operator.  The multi-dimensional interpolation in
Eq.~\ref{eqn:interp} needs to be performed many times, so computational
efficiency is critical.  In practice, a simple tri-linear interpolation is
suitably fast, but introduces an undesirable level of inaccuracy
(particularly in the [Fe/H] dimension).  The interpolation quality can be
empirically estimated by performing the operation in Eq.~\ref{eqn:interp}
across a $\{\vt_{\ast}\}^{\rm grid}$ grid point, and then comparing the
interpolated spectrum with the corresponding library spectrum. After an
extensive exploration of such calculations \citep[see also][]{husser12}, we
concluded that the best combination of speed and accuracy can be achieved by
pre-computing a {\it refined} spectral library using a cubic spline
interpolation with a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex,
0.1\,dex], and then performing tri-linear interpolation over that refined
grid.  Overall, this interpolation technique is found to be accurate within
a few percent per full-resolution model pixel and $\lesssim 1$\% once the
spectrum is downsampled to the pixels of the detector.  Ideally, this
pre-interpolation could be avoided if the spectral library was computed
over a refined grid (with a substantial up-front computational investment);
but for the time being, we can empirically propagate these interpolation
uncertainties into the likelihood calculations (as will be described in
Section~\ref{subsec:likelihood}).

@@

We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
spectral or flux dimensions.  For the former, we consider three kernels that contribute to the 
line-of-sight velocity distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, 
treats the instrumental spectral broadening.\footnote{For computational efficiency, we first 
pre-convolve the entire raw spectral library with a slight under-prediction of $\mathcal{F}_v^{\rm 
inst}$ and resample the spectra to a lower resolution.  Because the instrumental point spread 
function acts as a low-pass filter, this procedure reduces the size of the library by a factor of 
10 or more while still preserving the final Fourier content of the spectra.  Then, when any given 
spectrum is post-processed, the actual $\mathcal{F}_v^{\rm inst}$ used is treated as a smaller 
``correction" to this pre-processing kernel.}  For illustrative purposes, we assume 
$\mathcal{F}_v^{\rm inst}$ is a Gaussian with mean $v = 0$ and constant width $\sigma_v$ at all 
$\lambda$, although more sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm 
rot}$, characterizes the broadening induced by (projected) stellar rotation, parameterized by 
$v\sin{i}$ as described by \citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm 
dop} = \delta(v-v_z)$, incorporates the radial velocity through a Doppler shift.  The model 
spectrum is modified by the parameters $[\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, 
using a convolution in velocity-space,\footnote{In practice, these convolutions are performed as 
multiplications in Fourier-space to better preserve spectral information \citep[cf.,][]{tonry79}; 
the mathematical formalism is presented for clarity.}
\begin{eqnarray} \label{eqn:broadening}
\flam(\vt_{\ast}, \sigma_v, v\sin{i}, v_z) &=& \flam(\vt_{\ast}) \otimes \varphi_v \\
                                           &=& \flam(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop}, \nonumber
\end{eqnarray} 
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation} \label{eqn:resampling}
\flam(\vt_{\ast},  \sigma_v, v\sin{i}, v_z) \mapsto \vM(\vt_{\ast},  \sigma_v, v\sin{i}, v_z),
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-element model vector $\vM$ (and $N_{\rm pix}$ is the number of pixels in the 
spectrum).  Figure \ref{fig:broadening} shows a (condensed) graphical representation 
of these post-processing steps.

@@

The procedure so far, summarized in Eq.~\ref{eqn:interp}--\ref{eqn:scaling}, is composed of 
straightforward operations demanded by practical astronomical and computing issues.  If the data 
were {\it perfectly} calibrated, we could proceed to a likelihood calculation that makes a direct 
comparison with $\vM(\vT)$.  However, that is unlikely to be the case.  The key concern is that an 
imperfect calibration might produce mismatches in the underlying shapes of the data and model.  
Although such mismatches presumably occur at a low-level over relatively broad wavelength scales, 
they still might represent a non-trivial likelihood contribution, and thereby bias our estimates of 
the physical parameters.  This concern is usually treated externally to any modeling procedure, most often 
by dividing the observed spectrum (and model spectrum) by a polynomial in $\lambda$.  But that 
``normalization" implicitly assumes that there is no relevant information content in the spectral 
shape; if $\vT$ also contributes on the mismatched scales, then adopting this approach will corrupt 
the inferences of these parameters.  Moreover, in practice this approach is limited, since defining 
an appropriate polynomial becomes difficult in cases where the spectral line density is high (e.g., 
molecular bands for cool stars).

We employ an analogous, but more rigorous, approach to deal with this issue, cast {\it internal} to 
the modeling framework to propagate the uncertainties introduced by additional degrees of freedom 
in the model.  Deviations in the spectral shape introduced by calibration residuals are treated as 
explicit contributions to the model spectrum, enabling an exploration of the distribution of 
possible ``tweaks" to the calibration that can be compartmentalized into a set of nuisance 
parameters and then marginalized.  This way, the inferences for the stellar parameters will 
properly account for the underlying uncertainty in the calibration of the spectral shape.  In 
practice, this is achieved by distorting segments of the model spectrum with polynomials, 
$\mathsf{P}$ \citep[e.g.,][]{eisenstein06,koleva09}.  For data that has $N_{\rm ord}$ spectral 
orders, each denoted with index $o$, the model spectrum can be decomposed into the set of orders
\begin{eqnarray} \label{eqn:chebyshev}
\vM(\vT, \cheb) &=& \{ \vM_o(\vT) \times \mathsf{P}_o \} \\
                &=& \{ \vM_o(\vT) \times \sum_n c_o^{(n)} \, T_o^{(n)} \}, \nonumber
\end{eqnarray}
and $T^{(n)}$ is the $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
coefficients are considered a set of nuisance (hyper-)parameters, $\cheb = [\{c_o^{(0)}, c_o^{(1)}, 
\ldots, c_o^{(n-1)} \}]$.  With judicious priors, we can ensure that the unintended treatment of 
real spectral features (e.g., molecular bands) as residual calibration artifacts is negligible.  
The lowest-degree (scaling) coefficient, $c^{(0)}$, is naturally degenerate with the solid angle, 
$\Omega$.  Therefore, we enforce an additional constraint that the mean of the polynomial is 
unity.  For data with a single spectral order, this means simply setting $c^{(0)} = 1$.  In the 
multiple order case, we assign $c^{(0)} = 1$ in an arbitrary order as an anchor, but permit the 
$c^{(0)}$ in other orders to be different.  It is worth noting that this formalism can, in 
principle, be extended to develop models for completely uncalibrated spectra, rather than the 
residual calibration mismatches described here (with a suitable relaxation of the priors on 
$\cheb$).  Figure \ref{fig:chebyshev} offers a practical demonstration of how these nuisance 
parameters are applied.

\begin{figure}[!b]
\begin{center}
  \includegraphics{figs/chebyshev.pdf}
  \figcaption{
    The spectrum at the overlap of two echelle orders (23 \& 24). ({\it panel 1}) The flux-calibrated dataset shows a slight discrepancy of $\lesssim 3$\% between orders. ({\it panel 2}) To account for this residual error in the flux calibration, we multiply the model spectrum by a Chebyshev polynomial, whose coefficients are parameters that are part of the model. In principle, we could divide the data by these polynomials to recover what the true flux-calibrated data should be ({\it panel 3}), but instead we multiply the model by the polynomials to recover the original discrepancy between orders ( {\it panel 4}). This has the advantage of preserving the original dataset as a fixed quantity and makes the model (Equation~\ref{eqn:chebyshev}) linear in the Chebyshev polynomial coefficients.
\label{fig:chebyshev}}
\end{center}
\end{figure}

@@

Astronomical spectrographs are designed such that the detector over-samples the instrumental 
line-spread function with at least a few pixels.  Therefore, adjacent pixels never record 
completely independent samples of the true spectrum.  In that case, a difference between an 
observed and modeled spectral feature will create a correlated residual that spans multiple 
pixels.  This can be demonstrated practically through the autocorrelation of $\vR$: a slight model 
mismatch will produce correlated residuals over a characteristic scale similar to the instrumental
broadening kernel width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of 
these correlated residuals in real data; a significant autocorrelation signal is found on a $\sim$5 
pixel scale, corresponding to the 6.8\,km s$^{-1}$ FWHM of $\mathcal{F}_v^{\rm inst}$.

>>>>

%\documentclass[preprint]{aastex} 
\documentclass[iop,floatfix,twocolappendix]{emulateapj}

@@

\newcommand{\vM}{\mathsf{M}}
\newcommand{\vD}{\mathsf{D}}
\newcommand{\vR}{\mathsf{R}}
\newcommand{\vC}{\mathsf{C}}
\newcommand{\fM}{ \vec{{\bm M}}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ \vec{{\bm D}}}
\newcommand{\fDi}{D_i}
\newcommand{\fR}{ {\bm R}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}
\newcommand{\matern}{Mat\'{e}rn}
\newcommand{\HK}{$\textrm{H}_2$O-K2}
\newcommand{\cc}[2]{c_{#2}^{(#1)}}

\newcommand{\flam}{f_\lambda}
\newcommand{\vt}{ {\bm \theta}}
\newcommand{\vT}{ {\bm \Theta}}
\newcommand{\vp}{ {\bm \phi}}
\newcommand{\vP}{ {\bm \Phi}}
\newcommand{\cheb}{ \vp_{\mathsf{P}}}
\newcommand{\chebi}[1]{ \vp_{\textrm{Cheb}_{#1}}}
\newcommand{\Cheb}{ \vP_{\textrm{Cheb}}}
\newcommand{\Chebi}[1]{ \vP_{\textrm{Cheb}_{\ne #1}}} 
\newcommand{\cov}{ \vp_{\mathsf{C}}}
\newcommand{\covi}[1]{ \vp_{\textrm{cov}_{#1}}} 
\newcommand{\Cov}{ \vP_{\textrm{cov}}}
\newcommand{\Covi}[1]{ \vP_{\textrm{cov}_{\ne #1}}}

\newcommand{\allParameters}{\vT} 
\newcommand{\nuisanceParameters}{\vP}

\newcommand{\KK}{\mathcal{K}}
\newcommand{\Kglobal}{\KK^{\textrm{g}}}
\newcommand{\Klocal}{\KK^l}

@@

\shorttitle{Robust Spectroscopic Inference with Imperfect Models}
\shortauthors{Czekala et al.}

@@

\title{Robust Spectroscopic Inference with Imperfect Models}

\author{Ian Czekala, Sean M.~Andrews, et al.}
\affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\email{iczekala@cfa.harvard.edu}

\begin{abstract}
We present a modular, extensible framework for the spectroscopic inference of 
physical parameters based on synthetic model spectra.  The subtraction of an 
imperfect model from a continuously sampled spectrum introduces covariance 
between adjacent datapoints (pixels) into the residual spectrum.  In the limit 
of high signal-to noise data with large spectral range that is common for 
stellar parameter estimation, that covariant structure can bias the parameter 
determinations.  We have designed a likelihood function formalism to account 
for the structure of the covariance matrix, utilizing the machinery of Gaussian 
process kernels.  We specifically address the common problem of mismatches in 
model spectral line strengths (with respect to data) due to intrinsic model 
imperfections (e.g., in the atomic or molecular data, or radiative transfer 
treatment) by developing a novel local covariance kernel framework that 
identifies and self-consistently downweights pathological spectral line 
``outliers."  By fitting multiple spectra in a hierarchical manner, these local 
kernels provide a mechanism to learn about and build data-driven corrections to 
synthetic model spectral libraries.  The application of this method, 
implemented as a freely available open source code, is demonstrated by fitting 
the high resolution optical ($V$-band) spectrum of WASP-14, an F5 dwarf with a 
transiting exoplanet, and the moderate resolution near-infrared ($K$-band) 
spectrum of Gliese~51, an M5 dwarf. 
\end{abstract}
\keywords{stars: fundamental parameters --- methods: data analysis --- methods: statistical}

@@

All astronomers recognize that spectroscopy offers a wealth of information that 
can help characterize the properties of the observing target.  In the context 
of stellar astrophysics, spectroscopy plays many fundamental roles.  The 
relative strengths and widths of stellar absorption lines provide access to key 
parameters like effective temperature ($T_{\rm eff}$) and surface gravity 
($\log g$), enabling model comparisons in the Hertzsprung-Russell diagram to 
estimate the masses and ages so crucial to understanding stellar evolution, as 
well as individual elemental (and molecular) abundances or the collective 
``metallicity" (typically parameterized as [Fe/H]), facilitating study of the 
chemical hallmarks of different stellar populations.  With sufficient spectral 
resolution, the velocity content of a spectrum can convey crucial information 
about stellar rotation ($v \sin i$) and kinematics (e.g., association with a 
cluster or companion through the radial velocity, $v_z$).  While many fields 
benefit from the spectroscopic measurements of these stellar properties, there 
is acute interest in the exoplanet community.  There, all estimates of the 
planet properties are made {\it relative} to the host properties (e.g., the 
planet-to-host mass or radius {\it ratio} is constrained with the radial 
velocity or transit techniques, respectively).  Moreover, essential clues to 
the planet formation process are encapsulated in the dependences of planet 
frequency on host mass \citep[e.g.,][]{johnson07,howard10} and metallicity 
\citep[e.g.,][]{fischer05,buchhave14}.

@@

There are various well-tested techniques being used to compare these models with observed spectra 
and thereby infer basic stellar parameters; here we highlight three examples to illustrate some key 
ideas.  First is a straightforward empirical approach that relies on distilling an information-rich 
subset of the data, usually in the form of spectral line equivalent widths and/or local continuum 
shapes.  A combined sequence of the ratios of these quantities are found to be especially sensitive 
to one key parameter in the models, and therefore can be used to quantify it \citep[e.g., {\tt 
MOOG};][]{sneden73,gray94,reid95,rojas-ayala10,rojas-ayala12}.  This ``indexing" approach has the 
advantage of being trivially fast, but each condensed relationship is only informative over a 
limited swath of parameter-space.  A second technique exploits the cross-correlation of a spectrum 
with a suite of model templates to optimize a set of parameters, usually with some weighting 
applied to specific spectral regions \citep[e.g., {\tt SPC};][]{buchhave12}.  In this case, the 
speed advantage is maintained (perhaps enhanced) and more data content is used (particularly in the 
spectral dimension), thereby achieving higher precision even for data with comparatively low 
sensitivity; the disadvantage is that the model quality and parameter inferences are assessed in an 
empirical, rather than probabilistic, framework.  A third approach employs a direct, pixel-by-pixel 
comparison between model and data, and is usually associated with a spectral synthesis back-end 
(e.g., {\tt SME}; \citealt{valenti96}).  This technique has the benefits of parametric flexibility 
(e.g., one can fit for arbitrary abundances or structures) and a proper inference framework 
\citep[usually a least-squares approach, although increasingly in a Bayesian format;][]{shkedy07,
schoenrich13}, but it is computationally expensive and can be biased by systematics 
\citep[e.g.,][]{mann13}; in practice, this often requires an effort focused on a pre-selected 
subsample of the data.

In this article, we design a flexible forward-modeling approach to the general spectroscopic 
inference problem in a Bayesian framework, building on the best aspects of the latter two methods 
highlighted above.  Employing a non-trivial covariance matrix parameterized by both global 
(stationary) and local (non-stationary) Gaussian process kernels, we are able to account for 
residual pixel-to-pixel correlations that arise when comparing observed spectra to intrinsically 
imperfect models.  This approach minimizes the potential for bias, efficiently propagates 
systematic uncertainties into the parameter inferences, and ultimately can serve as tool for using
observations to guide substantial improvements in the models themselves.  An overview of the 
methodology of this approach is provided in Section \ref{sec:method}, including the covariance 
matrix formalism.  Some tests and example applications (for a high resolution optical spectrum of 
an F star, and a medium-resolution near-infrared spectrum of a mid-M star) of the method are 
described in Section \ref{sec:examples}.  Finally, a discussion of the potential utility of the 
technique, and the possibility of extending it to develop data-driven spectral models, is provided 
in Section \ref{sec:discussion}.  \\

@@

We opt for the computationally expedient approach that employs a library of model spectra, 
$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, 
[{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
{\it any} ``back-end" that generates a model spectrum.  In our adopted approach, the model spectrum 
for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library,
\begin{equation} \label{eqn:interp} 
\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}), 
\end{equation} 
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces a 
non-negligible level of uncertainty.  These interpolation uncertainties can be empirically 
estimated by performing the operation in Eq.~\ref{eqn:interp} across a $\{\vt_{\ast}\}^{\rm grid}$ 
grid point, and then comparing the interpolated spectrum with the corresponding library spectrum 
\citep[see also][]{husser12}.  We have developed this approach to propagate the interpolation 
uncertainties into the likelihood calculations, as described briefly in 
Section~\ref{subsec:likelihood} and elaborated in the Appendix.

@@

We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
spectral or flux dimensions.  For the former, we consider three kernels that contribute to the 
line-of-sight velocity distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, 
treats the instrumental spectral broadening.\footnote{For computational efficiency, we first 
pre-convolve the entire raw spectral library with a slight under-prediction of $\mathcal{F}_v^{\rm 
inst}$ and resample the spectra to a lower resolution.  Because the instrumental point spread 
function acts as a low-pass filter, this procedure reduces the size of the library by a factor of 
10 or more while still preserving the final Fourier content of the spectra.  Then, when any given 
spectrum is post-processed, the actual $\mathcal{F}_v^{\rm inst}$ used is treated as a smaller 
``correction" to this pre-processing kernel.}  For illustrative purposes, we assume 
$\mathcal{F}_v^{\rm inst}$ is a Gaussian with mean $v = 0$ and constant width $\sigma_v$ at all 
$\lambda$, although more sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm 
rot}$, characterizes the broadening induced by (projected) stellar rotation, parameterized by 
$v\sin{i}$ as described by \citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm 
dop} = \delta(v-v_z)$, incorporates the radial velocity through a Doppler shift.  The model 
spectrum is modified by the parameters $[\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, 
using a convolution in velocity-space,\footnote{In practice, these convolutions are performed as 
multiplications in Fourier-space to better preserve spectral information \citep[cf.,][]{tonry79}; 
the mathematical formalism is presented for clarity.}
\begin{eqnarray} \label{eqn:broadening}
\flam(\vt_{\ast}, \sigma_v, v\sin{i}, v_z) &=& \flam(\vt_{\ast}) \otimes \varphi_v \\
                                           &=& \flam(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop}, \nonumber
\end{eqnarray} 
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation} \label{eqn:resampling}
\flam(\vt_{\ast},  \sigma_v, v\sin{i}, v_z) \mapsto \vM(\vt_{\ast},  \sigma_v, v\sin{i}, v_z),
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-element model vector $\vM$ ($N_{\rm pix}$ is the number of pixels in the spectrum).  
Figure \ref{fig:broadening} shows a (condensed) graphical representation of these post-processing 
steps.

@@

% NEEDS UPDATED FIGURE
\begin{figure}[b!]
\includegraphics{figs/chebyshev.pdf}
\vspace{-0.75cm}
\figcaption{A demonstration of our treatment for residual calibration mismatches.  The observed 
spectra at the overlap of two echelle orders ({\it top}) have slightly ($\sim$1--3\%) discrepant 
shapes.  Chebyshev polynomials ({\it middle, top}) are designed to account for that mismatch ({\it 
middle, bottom}); rather, in practice, the model is scaled (multiplied) by those polynomials ({\it 
bottom}).  \label{fig:chebyshev}}
\end{figure}

The procedure so far, summarized in Eq.~\ref{eqn:interp}--\ref{eqn:scaling}, is composed of 
straightforward operations demanded by practical astronomical and computing issues.  If the data 
were {\it perfectly} calibrated, we could proceed to a likelihood calculation that makes a direct 
comparison with $\vM(\vT)$.  However, that is unlikely to be the case.  The key concern is that an 
imperfect calibration might produce mismatches in the underlying shapes of the data and model.  
Although such mismatches presumably occur at a low-level over relatively broad wavelength scales, 
they still might represent a non-trivial likelihood contribution, and thereby bias our estimates of 
the physical parameters.  This issue is usually treated externally to any modeling procedure, most 
often by dividing the observed spectrum (and model spectrum) by a polynomial in $\lambda$.  But 
that ``normalization" implicitly assumes that there is no relevant information content in the 
spectral shape; if $\vT$ also contributes on the mismatched scales, then adopting this approach 
will corrupt the inferences of these parameters.  Moreover, in practice this approach is limited, 
since defining an appropriate polynomial becomes difficult in cases where the spectral line density 
is high (e.g., molecular bands for cool stars).

We employ an analogous, but more rigorous, approach to deal with this issue, cast {\it internal} to 
the modeling framework to propagate the uncertainties introduced by additional degrees of freedom 
in the model.  Deviations in the spectral shape introduced by calibration residuals are treated as 
explicit contributions to the model spectrum, enabling an exploration of the distribution of 
possible ``tweaks" to the calibration that can be compartmentalized into a set of nuisance 
parameters and then marginalized.  This way, the inferences for the stellar parameters will 
properly account for the underlying uncertainty in the calibration of the spectral shape.  In 
practice, this is achieved by distorting segments of the model spectrum with polynomials, 
$\mathsf{P}$ \citep[e.g.,][]{eisenstein06,koleva09}.  For data that has $N_{\rm ord}$ spectral 
orders, each denoted with index $o$, the model spectrum can be decomposed into the set of orders
\begin{eqnarray} \label{eqn:chebyshev}
\vM(\vT, \cheb) &=& \{ \vM_o(\vT) \times \mathsf{P}_o \} \\
                &=& \{ \vM_o(\vT) \times \sum_n c_o^{(n)} \, T_o^{(n)} \}, \nonumber
\end{eqnarray}
where $T^{(n)}$ is the $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
coefficients are considered a set of nuisance (hyper-)parameters, $\cheb = [\{c_o^{(0)}, c_o^{(1)}, 
\ldots, c_o^{(n-1)} \}]$.  With judicious priors, we can ensure that the unintended treatment of 
real spectral features (e.g., molecular bands) as residual calibration artifacts is negligible.  
The lowest-degree (scaling) coefficient, $c^{(0)}$, is naturally degenerate with the solid angle, 
$\Omega$.  Therefore, we enforce an additional constraint that the mean of the polynomial is 
unity.  For data with a single spectral order, this means simply setting $c^{(0)} = 1$.  In the 
multiple order case, we assign $c^{(0)} = 1$ in an arbitrary order as an anchor, but permit the 
$c^{(0)}$ in other orders to be different.  Figure \ref{fig:chebyshev} offers a practical 
demonstration of how these nuisance parameters are applied to the model.

@@

Astronomical spectrographs are designed such that the detector over-samples the instrumental 
line-spread function with at least a few pixels.  Therefore, adjacent pixels never record 
independent samples of the true spectrum.  In that case, a difference between an observed and 
modeled spectral feature will create a correlated residual that spans multiple pixels.  This can be 
demonstrated practically through the autocorrelation of $\vR$: a slight model mismatch will produce 
correlated residuals over a characteristic scale similar to the instrumental broadening kernel 
width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of these correlated 
residuals in real data; a significant autocorrelation signal is found on a $\sim$5 pixel scale, 
corresponding to the 6.8\,km s$^{-1}$ FWHM of $\mathcal{F}_v^{\rm inst}$.

\begin{figure}[!htb]
\begin{center}
  \includegraphics{residuals_23.pdf}
  \includegraphics{class0_residuals.pdf}
  \includegraphics{class0_autocorrelation.pdf}
  \figcaption{({\it top}) A comparison of the data and a typical model with parameters drawn from 
the posterior distribution, along with the corresponding residual spectrum.  ({\it middle}) A 
zoomed view of the gray band in the top panels, highlighting the mildly covariant residual 
structure that is produced by slight mismatches between the data and model spectra.  ({\it bottom}) 
The autocorrelation of the residual spectrum.  Notice the substantial autocorrelation signal for 
offsets of $\lesssim 8$ pixels, demonstrating clearly that the residuals are not described well by 
white (Poisson) noise alone. \label{fig:class0}}
\end{center}
\end{figure}
